Multilevel selection in nonergodic systems: the evolutionary advantage of cooperation and especialization.

### 2021-11-11 talk/transitions.tex
Talk

\centering
\tikz{
    \node[latent] (e) {$e$};
    \node[const, right=of e] (en) {\ $p(e)=\text{Beta}(1,1)$};
    \node[const, left=of e] (ne) {Estrategias: \ \ \ };

    \node[latent, below=of e] (r) {$a$};
    \node[const, right=of r] (rn) {$P(a|e,T) = \text{Binomial}(a|T,e)$};
    \node[const, left=of r] (nr) {Aptitud: \ \ \ };

@@

\vspace{0.6cm}
\pause

@@

@@

>>>>

\begin{frame}[plain]
\begin{textblock}{160}(0,4)
 \centering \LARGE
\en{Base model}\es{Modelo base} $M_0$
\end{textblock}
\vspace{1.3cm}

\centering
\tikz{
    \node[latent] (e) {$e$};
    \node[const, right=of e, xshift=0.2cm] (en) {\ $p(e)=\text{Beta}(1,1)$};
    \node[const, left=of e, xshift=-0.4cm] (ne) {Estrategias: \ \ \ };

    \node[latent, below=of e] (r) {$a^t$};
    \node[const, right=of r, xshift=0.2cm] (rn) {$P(a|e) = \text{Bernoulli}(a|e)$};
    \node[const, left=of r, xshift=-0.4cm] (nr) {Aptitud: \ \ \ };

    \plate {observaciones} {(r)} {$ 1 \leq t \leq T$};

@@

\vspace{0.2cm}
\pause

@@

\vspace{-0.7cm}
\pause

\begin{equation*}
P(a(1\dots T), e | M_0) = \prod_t^T P(a(t)|e) \approx  P(a=1|e)^{p^*} P(a=0|e)^{1-p^*}
\end{equation*}

\begin{frame}[plain]
\begin{textblock}{160}(0,4)
 \centering \LARGE
\en{General model}\es{Modelo general} $M_{e,n}$
\end{textblock}
\vspace{1.5cm} \centering

\tikz{

    \node[latent, minimum size=0.9cm] (e1) {$s^{t}$};
    \node[latent, right=of e1, minimum size=0.9cm] (e2) {$s^{t+1}$};
    \node[const, right=of e2] (e2n) {$P(s^{t+1}|s^{t}) = \frac{1}{n}$};
    \node[const, right=of e2, yshift=-1cm] (e2n) {$P(s^{0}) = \text{Prior}$};
    \node[const, right=of e2, yshift=-2cm] (e2n) {$P(a_s^t|e) = \text{Bernoulli}(a_s^t|e)$};

    \node[latent, below=of e1, minimum size=0.9cm] (a) {$a^{t}$};
    \edge {e1} {a,e2};
    }

\begin{equation*}
\begin{split}
\onslide<2->{P(s^{t+1}, a^t | M_{n,e} ) &=\frac{1}{3} \text{Prior}^t  \overbrace{P(a_1^t|e)}^{\hfrac{\text{Likelohood}}{\text{individual 1}}} + \frac{1}{3} \text{Prior}^t  \overbrace{P(a_2^t|e)}^{\hfrac{\text{Likelohood}}{\text{individual 2}}}  + \frac{1}{3} \text{Prior}^t  \overbrace{P(a_3^t|e)}^{\hfrac{\text{Likelohood}}{\text{individual 3}}}}  \\
\onslide<3->{&= \text{Prior}^t \underbrace{(\frac{1}{3} P(a_1^t|e) + \frac{1}{3} P(a_2^t|e) + \frac{1}{3} P(a_3^t|e))}_{f(a|e,n=3)\text{ fitness cooperativo}} = \text{Prior}^{t+1}}
\end{split}
\end{equation*}

\begin{frame}[plain]
\begin{textblock}{160}(0,4)
 \centering \LARGE
\en{General model}\es{Modelo general} $M_{e,n}$
\end{textblock}
\vspace{1cm} \centering

%
\only<1>{
\vspace{0.5cm}
\begin{equation*}
f(a|e,n=3) =
\begin{cases}
(1-e) & \text{ si } \texttt{sum}(a) = 0 \\
\frac{1}{3} e + \frac{2}{3} (1-e)  & \text{ si } \texttt{sum}(a) = 1 \\
\frac{2}{3} e + \frac{1}{3} (1-e)    & \text{ si } \texttt{sum}(a) = 2 \\
e & \text{ si } \texttt{sum}(a) = 3
\end{cases}
\end{equation*}
%
}

\onslide<2->{
\begin{equation*}\label{eq:fitness_cooperador}
f(a|e,n) = \frac{x}{n} e + \frac{n-x}{n}(1-e) \ \ \ \ \ \ \ \ \text{ con } x = \texttt{sum}(a)
\end{equation*}
}

\vspace{-0.8cm}

\onslide<3->{
\begin{equation*}
P(s^{T+1}, \bm{a}|e,n) = \text{Prior}^0 \prod^T_{t=1} f(a^t|e,n) \onslide<4->{\approx \overline{f}_{e,n}^T}
\end{equation*}
}

\vspace{-0.8cm}

\onslide<5->{
\begin{equation*}
\overline{f}_{e,n} = \prod_{a=0}^n f(a|e,n)^{\text{Binomial}(x|n,p^*)}
\end{equation*}
}

\vspace{-0.8cm}

\onslide<6->{
\begin{equation*}
\lim_{n \rightarrow \infty} \overline{f}_{e,n} =  p^* \, e + (1 - p^*) \, (1-e)
\end{equation*}
}

\begin{frame}[plain]
\begin{textblock}{160}(0,4)
 \centering \LARGE
\en{General model}\es{Modelo general} $M_{e,n}$
\end{textblock}
\vspace{1.5cm} \centering

\includegraphics[width=0.66\textwidth]{../figures/pdf/tasa-temporal-0.pdf} \hspace{1cm}

\begin{frame}[plain]
\begin{textblock}{160}(0,4)
 \centering \LARGE
\en{General model}\es{Modelo general} $M_{e,n}$
\end{textblock}
\vspace{1.5cm} \centering

\includegraphics[width=0.66\textwidth]{../figures/pdf/tasa-temporal-1.pdf} \hspace{1cm}

\begin{frame}[plain]
\begin{textblock}{160}(0,4)
 \centering \LARGE
\en{General model}\es{Modelo general} $M_{e,n}$
\end{textblock}
\vspace{1.5cm} \centering

\includegraphics[width=0.66\textwidth]{../figures/pdf/tasa-temporal-2.pdf} \hspace{1cm}

\begin{frame}[plain]
\begin{textblock}{160}(0,4)
 \centering \LARGE
\en{General mixed model}\es{Modelo general mixto} $M_{e,n,N}$
\end{textblock}
\vspace{1.3cm} \centering

\tikz{    
    \node[latent, minimum size=0.9cm] (e1) {$s^{t}$};
    \node[latent, right=of e1, minimum size=0.9cm] (e2) {$s^{t+1}$};
    \node[const, right=of e2] (e2n) {$P(s^{t+1}|s^{t}) = \begin{cases}\frac{1}{N}  & s^t \leq n \\ \mathbb{I}(s^{t+1}=s^t) & s^t > n\end{cases}$};
    \node[const, right=of e2, yshift=-1cm] (e2n) {$P(s^{0}) = \text{Prior}$};
    \node[const, right=of e2, yshift=-2cm] (e2n) {$P(a_s^t|e) = \text{Bernoulli}(a_s^t|e)$};

    \node[latent, below=of e1, minimum size=0.9cm] (a) {$a^{t}$};
    \edge {e1} {a,e2};
}

\pause
\vspace{0.4cm}

\begin{align*}
\centering
P(M|e=0.71,N=2) = \begin{tabular}{|c|c|c|c|}
        \hline
        $n=0$ & $n=1$ & $n=2$ \\ \hline
        $1/4$ & $1/2$ & $1/4$ \\ \hline
\end{tabular}
\end{align*}

\begin{frame}[plain]
\begin{textblock}{160}(0,4)
 \centering \LARGE
\en{Model selection}\es{Selección de modelo} $M_{e,n,N}$
\end{textblock}
\vspace{1.5cm} \centering

\includegraphics[width=0.66\textwidth]{../figures/pdf/multilevel-selection-6.pdf}\hspace{1cm}

@@

\begin{frame}[plain]
\begin{textblock}{160}(0,4)
 \centering \LARGE
\en{Growth rate for mixed populations}\es{Tasa de crecimiento para poblaciones mixtas} $M_{e,n,N}$
\end{textblock}
\vspace{1.3cm} \centering

\begin{equation*}
P(s=1,a|e,n,N) = \frac{x}{N} e + \frac{n-x}{N}(1-e) = \frac{n}{N} f(a|e,n)
\end{equation*}

\vspace{0.5cm}

\begin{equation*}
P(s=N, \bm{a}^{:t}|e,n,N) =
\begin{cases}
 P(a_s^1|e) & \ \  t=1 \\
 (P(\bm{a}^{:t-1},s=N|e,n,N) + P(\bm{a}^{:t-1}, s=1|e,n,N)) p(a_s^t|e) & \ \  t>1
\end{cases}
\end{equation*}

\begin{frame}[plain]
\begin{textblock}{160}(0,4)
 \centering \LARGE
\en{Growth rate for mixed populations}\es{Tasa de crecimiento para poblaciones mixtas} $M_{e,n,N}$
\end{textblock}
\vspace{1.3cm} \centering

\includegraphics[width=0.66\textwidth]{../figures/pdf/multilevel-selection-5.pdf}

\begin{frame}[plain]
\begin{textblock}{160}(0,4)
 \centering \LARGE
\en{Fitness for mixed populations}\es{Fitness para poblaciones mixtas} $M_{e,n,N}$
\end{textblock}
\vspace{1.3cm} \centering

\includegraphics[width=0.66\textwidth]{../figures/pdf/multilevel-selection-7.pdf}

\begin{frame}[plain]
\begin{textblock}{160}(0,4)
 \centering \LARGE
\en{Fitness for mixed populations}\es{Fitness para poblaciones mixtas} $M_{e,n,N}$
\end{textblock}
\vspace{1.3cm} \centering

\begin{figure}[H]
    \centering
    \begin{subfigure}[b]{0.32\textwidth}
    \includegraphics[width=\linewidth]{../figures/pdf/multilevel-selection-1.pdf}
    \caption{9/9}
    \label{fig:multilevel-selection-1}
    \end{subfigure}
    \begin{subfigure}[b]{0.32\textwidth}
    \includegraphics[width=\linewidth]{../figures/pdf/multilevel-selection-2.pdf}
    \caption{8/9}
    \label{fig:multilevel-selection-2}
    \end{subfigure}
    \begin{subfigure}[b]{0.32\textwidth}
    \includegraphics[width=\linewidth]{../figures/pdf/multilevel-selection-3.pdf}
    \caption{7/9}
    \label{fig:multilevel-selection-3}
    \end{subfigure}
\end{figure}
%
